---
title: "Untitled"
author: 
date:
output: html_document
---

```{r}

library(randomForest)
library(tidyverse)
library(grid)
library(mvtnorm)

```

Here we create simulated data used to test the code. 
```{r}

#n is number of observations in dataset
n <- 1000
#p is number of variables
p <- 5
#our weights are given by wt_vec <- c(5,5,1,0,0)
err <- rnorm(n, mean = 0, sd = 0.05)
siml.data <- as.data.frame(matrix(rnorm(n*p, mean = 0, sd = 1), nrow = n))
siml.data <- siml.data %>% 
  mutate(Y = 5*V1+5*V2+V3+err)


```


Next want to write a function which takes a dataframe and applys the added-variable plot then permute scheme to each predictor in the dataset. 

```{r}

#first need to take our data frame and make the p+1 datasets that we apply the random forest to

#itr_col takes as input an integer and removes that column from the dataset
#note: should make itr_col more robust
itr_col <- function(i, data){
  select(data,-i)
}

#df_combs is a function which takes as input a dataframe and returns all iterations of the
#dataframe where one predictor has been removed. Output is a p+1 element list where each element
#is a dataframe with the ith variable removed. the last element of the list is the entire dataframe
df_combs <- function(data){
  p <- ncol(data)-1
  df.list <- map(1:p, itr_col, data = data)
  df.list[[p+1]] <- data
  df.list
}

#extract_rf_pred is a function which takes as input a list of randomforest objects, an index value, 
#and the number of predictors in the model and outputs the predicted values of the random forest in 
#a dataframe. extract_rf_pred is mainly for use in map_dfc, which binds by column the predicted values in 
#data.list as a dataframe. Primarily for use in each_pred_rf
extract_rf_pred <- function(i, data.list, p){
 new.df <- as.data.frame(data.list[[i]]$predicted) 
 colnames(new.df) <- ifelse(i <= p, paste("PredWoVar", as.character(i), sep = ""),
                            "PredFullMod")
 new.df
}

#each_pred_rf is a function whose input is a list of dataframes that come from output of df_combs and runs the 
#randomForest function on each dataframe using the map function. Output is dataframe of predicted values along with actual value of Y as last column. Second to last column is the predicted value of full model. 
each_pred_rf <- function(data.list, ntree1, replace = TRUE){
  p <- length(df.list)-1
  rf.list <- map(.x = data.list, function(x) 
    randomForest(Y~., data = x, ntree = ntree1, replace = replace, importance = TRUE))
  rf.df <- map_dfc(1:(p+1), extract_rf_pred, data.list = rf.list, p = p)
  Y <- data.list[[1]]$Y
  new.df <- cbind(rf.df, Y)
  imp <- importance(rf.list[[6]])
  list(new.df, imp)
}



#extract_add_var takes as input an index value and data frame and 
#outputs a dataframe of the basic added variable data frame where 
#x.res is the difference between predicted values of full model and 
#predicted values of model with out jth variable
#y.res is the residual of Y and predicted values of model without jth variable.
#For use with map in rf_add_var
extract_add_var <-function(i, df){
  PredFullMod <- as.name("PredFullMod")
  Y <- as.name("Y")
  V <- as.name(paste("PredWoVar", as.character(i), sep =""))
  x.res <- df[[PredFullMod]]-df[[V]]
  y.res <- df[[Y]]-df[[V]]
  new.df <- data.frame(x.res, y.res)
  colnames(new.df) <- c(paste("diff.full.woVar", as.character(i), sep = ""), "y.res")
  new.df
}

#rf_add_var takes as input the output of each_pred_rf and outputs a list of length p 
#in which each entry is a data frame corresponding to an added variable plot for the jth predictor in the 
#data set.
rf_add_var <- function(data.list){
  rf.df <- data.list[[1]]
  p <- length(rf.df)-2
  add.var.list <- map(1:p, extract_add_var, df = rf.df)
  add.var.list
}

#rf_add_var_imp takes as input a list of add_var df's from rf_add_var and runs
#a random forest on the y-residuals with x-residuals as input.
#output is a list of random forest objects. Might change output to be just variable importance values. 
rf_add_var_imp <- function(data.list, ntree2, replace = TRUE){
  p <- length(data.list)
  rf.add.imp.list <- map(.x = data.list, function(x) 
    randomForest(y.res~., data = x, ntree = ntree2, replace = replace, importance = TRUE))
  rf.add.imp.list
}

extract_imp <-function(i, data.list){
  new.df <- as.data.frame(t(importance(data.list[[i]])))
  new.df
}

extract_var_imp <- function(data.list){
  p <- length(data.list)
  new.df <- map_dfc(1:p, extract_imp, data.list = data.list)
  rownames(new.df) <- c("%IncMSE", "IncNodePurity")
  as.data.frame(t(new.df))
}


#add_var_randomforest is a wrapper for the previous functions
add_var_randomforest <- function(data, ntree1, ntree2, replace = TRUE){
  #to get the copy of the data with one predictor removed
  df.list <- df_combs(data)
  #running initial randomFoest on each data frame
  rf.list <- each_pred_rf(df.list, ntree1 = ntree1)
  #tidying the data to compute added variable plots
  rf.add.var.list <- rf_add_var(rf.list)
  #running randomForests on each plot
  rf.add.imp.list <- rf_add_var_imp(rf.add.var.list, ntree2 = ntree2)
  #extracts variable importance values for each randomForest ran on added variable plots
  add.var.imp <- extract_var_imp(rf.add.imp.list)
  #output is list containing data frame of added variable plot variable importances, 
  #rf.add.var.list which are dataframe for added variable plots, and 
  #the variable importances from full model run in each_pred_rf
  list(add.var.imp, rf.add.var.list, rf.list[[2]])
}

#input for rf_added_var_plot is output of 
#add_var_randomforest. Output is plot of added variable plots for 
#the random forest arranged in a grid.

plot_add_var <- function(i, df.list){
  df <- df.list[[i]]
  x.name <- colnames(df)[1]
  y.name <- colnames(df)[2]
  plot.obj <- ggplot(df, aes_string(x = x.name, y = y.name))+geom_point()
  plot.obj
}

rf_added_var_plot <- function(data.list){
    df.list <- data.list[[2]]
    p <- length(df.list)
    gg.list <- map(1:p, plot_add_var, df.list = df.list)
    nCol <- floor(sqrt(p))
    do.call("grid.arrange", c(gg.list, ncol = nCol))
}

#Now to test
rf.add.var.imp <- add_var_randomforest(siml.data, ntree1 = 1000, ntree2 = 1000, replace = FALSE)
rf.add.var.imp[[1]]
rf.add.var.imp[[3]]
rf_added_var_plot(rf.add.var.imp)


```

Now ready to run simulations. First and second simulated datasets are from Strobl (2008). 
First simulated dataset is Strobl's with iid variables.
Second simulated dataset is Strobl's with correlated variables.
Both are cases of a response in which the effect of added variables is linear. 

```{r}

set.seed(17)
#iid
#n1 is number of observations in dataset
n1 <- 2000
#p1 is number of variables
p1 <- 12
#our weights are given by wt_vec <- c(5,5,2,0,-5,-5,-2,0,0,0,0,0) 
err1 <- rnorm(n1, mean = 0, sd = 0.5)
siml.data1 <- as.data.frame(matrix(rnorm(n1*p1, mean = 0, sd = 1), nrow = n1))
siml.data1 <- siml.data1 %>% 
  mutate(Y = 5*V1+5*V2+2*V3+-5*V5+-5*V6+-2*V7+err1)

#correlated
#n2 is number of observations in dataset
n2 <- 2000
#p2 is number of variables
#p2 <- 12
#our weights are given by wt_vec <- c(5,5,2,0,-5,-5,-2,0,0,0,0,0) 
#covariance matrix
v1 <- c(1,0.9,0.9,0.9)
v2 <- c(0.9,1,0.9,0.9)
v3 <- c(0.9,0.9,1,0.9)
v4 <- c(0.9,0.9,0.9,1)
w <- rep(0, times = 8)
u1 <- c(rep(0, times = 4), 1, rep(0, times = 7))
u2 <- c(rep(0, times = 5), 1, rep(0, times = 6))
u3 <- c(rep(0, times = 6), 1, rep(0, times = 5))
u4 <- c(rep(0, times = 7), 1, rep(0, times = 4))
u5 <- c(rep(0, times = 8), 1, rep(0, times = 3))
u6 <- c(rep(0, times = 9), 1, 0,0)
u7 <- c(rep(0, times = 10), 1, 0)
u8 <- c(rep(0, times = 11), 1)

Sigma2 <- matrix(c(v1,w,v2,w,v3,w,v4,w,
                   u1,u2,u3,u4,u5,u6,u7,u8), nrow = 12, ncol = 12)
err2 <- rnorm(n2, mean = 0, sd = 0.5)
siml.data2 <- as.data.frame(rmvnorm(n = n2, mean = rep(0, times = 12), sigma = Sigma2))
siml.data2 <- siml.data2 %>% 
  mutate(Y = 5*V1+5*V2+2*V3+-5*V5+-5*V6+-2*V7+err2)

#run the simulations with and without replacement for bootstrapping
siml1.wrep.results <- add_var_randomforest(data = siml.data1, ntree1 = 2000, ntree2 = 2000, replace = TRUE)
siml1.worep.results <- add_var_randomforest(data = siml.data1, ntree1 = 2000, ntree2 = 2000, replacee = FALSE)
siml2.wrep.rseults <- add_var_randomforest(data = siml.data2, ntree1 = 2000, ntree2 = 2000, replace = TRUE)
siml2.worep.results <- add_var_randomforest(data = siml.data2, ntree1 = 2000, ntree2 = 2000, replace = FALSE)



```

